---
title: "Exercices in linear models and experimental design"
author: "Nuria Garcia"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    toc: no
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, cache = FALSE,
                      echo = FALSE, warning = FALSE, message = FALSE)
```
# Introduction

These exercises describe experimental situations that have been studied using some omics technology like microarrays.

# Comparison between three types of breast cancer

This case study is based on a paer published in  http://www.ncbi.nlm.nih.gov/pubmed/15897907 whose data are available in GEO as series GSE1561 series on the following link
http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE1561

The researchers investigated three types of breast
cancer tumors: apocrine (APO), basal (BAS) and luminal (LUMI). The classification is based on the resistance of tumors to estrogen and androgen receptors.

- Tumors classified as "APO" are negative for estrogen receptor (ER-) and positive for the androgen receptor (AR +).
- Those classified as "LUMI" are ER + and AR + and
- Those classified as "BAS" are ER- and AR.

The assignment of each sample to an experimental group can be obtained from this link:
http://www.ncbi.nlm.nih.gov/geo/gds/profileGraph.cgi?gds=1329

Obviously this is an observational study but its analysis can be done using a linear model approach as well.

1. Identify the experimental factors and their levels.
2. Write down the design matrix associated with this study design.
3. Build the contrast matrix needed to compare each tumor type with the oher two, that is:

    1. "APO" vs “LUMI”
    2. “APO" vs “BAS”
    3. “LUMI" vas "BAS"
    

```{r readTargets}
require (limma)
targetsLocation<- read.csv("taget.txt")
targetsLocation
#targets<- read.delim(targetsLocation, row.names=1)
```


```{r designMatrix}
design<-matrix(
  c(1,0,1,0,1,0,0,0,0,0,0,0,0,1,0,1,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,0,1,0,1,0,0,0,0,0,0,0,0,1,0,1,0,1),nrow=12,byrow=F)
#design2 <-model.matrix(~ 0+targets$Group) # We need to put a 0 to not have an intercept line
colnames(design)<- colnames(design)<-c("miR_B1", "miR_B2", "Scr_B1", "Scr_B2")
#rownames(design) <- rownames(design)<- targets$SampleIDs 
print(design)
```

```{r contrastsMatrix}
cont.matrix <- makeContrasts (
  m1vsm2 = miR_B1-miR_B2, #this must be the column names predetarminated before.
  m1vss1 = miR_B1-Scr_B1,
  m1vss2 = miR_B1-Scr_B2,
  m2vss1 = miR_B2-Scr_B1,
  m2vss2 = miR_B2-Scr_B2,
  s1vss2 = miR_B1-Scr_B2,
  
  levels=design)
```


