---
title: "Exploratory analysis of a GEO dataset"
author: "Núria Garcia Deltiempo"
date: "15/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:
The goal of this practical is to learn how to work with Markdown and Bioconductor and understand the steps that we follow to analyze our data. 

# Summary:

This study titled "The effect of miRNA-135b overexpression on the gene expression profile of LNCaP cells" is about MicroRNAs (miRNAs) regulate a wide range of cellular signaling pathways and biological processes in both physiological and pathological states such as cancer. We have previously identified miR-135b as a direct regulator of androgen receptor (AR) protein level in prostate cancer (PCa). We wanted to further explore the relationship of miR-135b to hormonal receptors, particularly estrogen receptor α (ERα). Here we show that miR-135b expression inversely correlates with ER protein in two independent breast cancer (BCa) patient cohorts (101 and 1302 samples) and with AR protein in 47 PCa patient samples. We identify ERα as a novel miR-135b target by demonstrating miR-135b binding to the 3’UTR of the ERα and decreased ERα protein and mRNA level in breast cancer cells upon miR-135b overexpression. miR-135b inhibits proliferation of hormone receptor positive cancer cell lines as shown by overexpression in ERα-positive BCa cells (MCF-7) and AR-positive PCa cells (LNCaP, 22Rv1) when grown in 2D. To identify other genes regulated by miR-135b we performed gene expression studies and found a potential link to the hypoxia-inducible factor-1α (HIF1α) pathway. We show that miR-135b influences the protein level of the inhibitor for hypoxia-inducible factor-1 (HIF1AN), which also demonstrated an inverse correlation with miR-135b in a cohort of breast tumor samples. Taken together, our study demonstrates that miR-135b regulates ERα, AR and HIF1AN protein levels and proliferation in ERα -positive breast and AR-positive-prostate cancer cells.

## Contributors:
Aakula A, Leivonen S, Hintsanen P, Aittokallio T, Ceder Y, Børresen-Dale A, Perälä M, Östling P, Kallioniemi O

# Data description:

Here we are introducing the data that we are goig to work with.
```{r readTarget}
target <- read.csv("/home/nuria/Escritorio/Garcia_Nuria_OmicsTechniques/Exercise_1/taget.txt", sep= "")
expresion <- as.matrix(read.delim("/home/nuria/Escritorio/Garcia_Nuria_OmicsTechniques/Exercise_1/expresion.txt", row.names = 1))
  ```


```{r}
dim(expresion)
summary(expresion)
boxplot(expresion)
```

Here we are installing the necessary package to handle the data that we have.

```{r}
require(Biobase)
```


```{r}
myEset <- ExpressionSet(expresion)
class(myEset)
show(myEset)

```

Here we are analyzing the covariates, to see how our data act.
```{r}
columnDesc <-  data.frame(labelDescription= c("Sample Names", "Group"))
myAnnotDF <- new("AnnotatedDataFrame", data=target, varMetadata= columnDesc)
show(myAnnotDF)
```

```{r}
phenoData(myEset) <- myAnnotDF
```

```{r}
rownames(pData(myAnnotDF))<-pData(myAnnotDF)$sampleName
myEset <- ExpressionSet(assayData=expresion, phenoData=myAnnotDF)
show(myEset)
```

```{r}
myEset <- ExpressionSet(assayData=expresion, 
                        phenoData=myAnnotDF, 
                        featureNames =myGenes)
```

Here is where we start storing the data.
```{r}
myDesc <- new("MIAME", name= "Anna Aakula",
            contact= "anna.aakula@fimm.fi" ,
            title="The effect of miRNA-135b overexpression on the gene expression profile of LNCaP cells")
print(myDesc)
```

```{r}
myEset <- ExpressionSet(assayData=expresion, 
                        phenoData=myAnnotDF,
                        fetureNames =myGenes,
                        experimentData = myDesc)
```

```{r}
dim(exprs(myEset))
class(phenoData(myEset))
class(pData(phenoData(myEset)))
head(pData(phenoData(myEset)))
head(pData(myEset))
```

Here we apply this functions to make data manipulation consistent.
```{r}
smallEset <- myEset[1:15,c(1:3,6:8)]
dim(exprs(smallEset))
dim(pData(smallEset))
head(pData(smallEset))
all(colnames(exprs(smallEset))==rownames(pData(smallEset)))
```

Here we are creating the dataset for the young patients, smaller than 30 years.
```{r}
youngEset <- myEset[,pData(myEset)$age<30]
dim(exprs(youngEset))
head(pData(youngEset))
```

```{r}
if (!require(GEOquery)) {
  BiocManager::install("GEOquery")
}
require(GEOquery)
gse <- getGEO("GSE57820")
class(gse)
names(gse)
gse[[1]]
esetFromGEO <- gse[[1]]


```

















